Data-based completion of grey-box models with neural networks

### 2020-09-22 systemID_NNs.tex
add things.

\title{System ID and control with NNs}
\author{Problem formulation}
\date{\today}

@@

\section{Process modeling}
Consider a nonlinear plant evolving as
\begin{align*}
  x^+ &= f(x, u) \\
  y &= h(x)
\end{align*}
in which $x$ is the plant state (vector), $u$ is the manipulated control input,
and $y$ is the measurement. Assume that we don't have any 
disturbances/measurement and process noise 
for now, and we'll add those after we 
understand the issues in this nominal case.

\subsection{Completion of the u-y model}
In this class of problems, we assume that grey-box modeling is performed 
for the plant, and we wish to augment the grey-box model with neural networks
for improvement in the accuracy of the model. From first-principles modeling, 
we have the following (in continuous time, as grey-box modeling will 
almost always be in continuous time).
\begin{align*}
  \dot{x}_g &= f_g(x_g, u) \\
  y &= h_g(x_g)
\end{align*}

From Taken's embedding theorem, we know that the original plant state
can be reconstructed using a history of measurements and control inputs.
Denote this history as 
\begin{align*}
  \mathbf{y}_{k:k-\tau} &= [y(k)', y(k-1)', ..., y(k-\tau)']' \\
  \mathbf{u}_{k-1:k-\tau} &= [u(k-1)', u(k-2)', ..., u(k-\tau)']'
\end{align*}

So from the embedding theorem for observable nonlinear plants, 
there exists a $\phi(\cdot)$ and number of delay 
time steps $\tau$, such that 
$x(k) = \phi(\mathbf{y}_{k:k-\tau}, \mathbf{u}_{k-1:k-\tau})$.

Now depending on the application, an approximation of $\phi(\cdot)$
with a neural network (say $\phi_{NN}(\cdot)$) 
can be used to augment the grey-box model 
in several ways. The following is proposed in the recent 
paper by Kevrekidis.
\begin{align*}
  x_g^+ &= f_g(x_g, u, 
  \theta(\phi_{NN}(\mathbf{y}_{k:k-\tau}, \mathbf{u}_{k-1:k-\tau})))\\
  y &= h_g(x_g)
\end{align*}

in which $\theta(\cdot)$ are a set of grey-box parameters which 
depend on the plant states. For example, rate constants which 
depend on temperature, etc. An RK4 discretization is performed
to get this model in discrete time.

A different type of problems could be say, there are unknown
side reactions in the plant we don't about, which are affecting the 
physical grey-box 
states. In these problems, we can augment the grey-box 
model with a NN as follows
\begin{align*}
x_g^+ &= f_g(x_g, u) + f_{NN}(\mathbf{y}_{k:k-\tau}, \mathbf{u}_{k-1:k-\tau})\\
y &= h_g(x_g) 
\end{align*}
in which, we are anticipating that $f_{NN}$,
a neural network can learn the difference between the 
true grey-box state dynamics and what is modeled with 
the function ($f_g(\cdot)$). We assume that the measurements
are directly some subset of the physical grey-box states.

\textbf{Need to discuss the following issues:}
\begin{itemize}
  \item What are the example problems/applications 
  where both kind of modeling 
  strategies make sense, and where should we start.
  Reproduce results from the Kevrekidis paper with measurement 
  noise?
  \item The model training step will be the optimization of 
  the following multi-step-ahead prediction-error
  \begin{equation*}
    \underset{W_i, b_i}{\textnormal{min}} \sum_{k=0}^{t-1} (\hat{y}(k) - y(k))^2
  \end{equation*}
  in which ($W_i, b_i$) are the weights of the NN. This optimization 
  is for one ID experiment, and requires an initial grey-box 
  state $x_g(0)$ to make the forecast. If we have data 
  for multiple ID experiments, we'll need initial states for 
  all the ID experiments. In theory, we can add the initial 
  states as decision variables in the optimization, but 
  unfortunately tensorflow 
  does not has a mechanism to do this.
\end{itemize}

\subsection{Adding feedforward disturbance forecasts}
In this class of problems, we assume that we have a pretty 
good u-y model from first-principles knowledge, and we'll use 
neural networks to make feedforward disturbance forecasts
for use in online MPC. HVAC is a pretty good application.

\section{Optimization in MPC with neural networks as model}
This is probably a late fall quarter or a winter break project. \\
a) Try Casadi/MPC tools. 
If tanh is the activation function (optimization is smooth),
this should be fine. \\
b) Try a global optimization solver. \\
c) Train another NN with solutions from a global optimization solver.

>>>>

\title{Data-based completion of grey-box models with neural networks}
\author{Problem formulation}
\date{\today}

@@

\section{Process modeling}
Consider a nonlinear plant evolving in continuous time as
\begin{align*}
  \dot x &= f_p(x, u) \\
  y &= C_px
\end{align*}
in which $x$ is the plant state (vector), $u$ is the manipulated control input,
and $y$ is the measurement. Assume that we don't have any 
disturbances/measurement and process noise 
for now, and we'll add those after we 
understand the issues in the nominal case.

\subsection{Completion of the u-y model}
In this class of problems, we assume that grey-box modeling is performed 
for the plant, and we wish to augment the grey-box model with neural networks
for improvement in the accuracy of the model. From first-principles modeling, 
we have the following (in continuous time, as grey-box modeling will 
almost always be in continuous time).
\begin{align*}
  \dot{x}_g &= f_g(x_g, u) \\
  y &= C_gx_g
\end{align*}

in which $x_g$ is some set of physical plant states chosen to be 
modeled. Denoted these physical states as $x_g=Cx$.
For observable nonlinear systems, 
we can construct the original plant state
using a history of measurements and control inputs.
Denote this history as 
\begin{align*}
  \mathbf{y}_{k:k-N} &= [y(k)', y(k-\Delta)', ..., y(k-N\Delta)']' \\
  \mathbf{u}_{k-1:k-N} &= [u(k-1)', u(k-2)', ..., u(k-N\Delta)']'
\end{align*}

There exists a $\phi(\cdot)$ and a number of delay 
time steps $N$, such that 
$x(k) = \phi(\mathbf{y}_{k:k-N}, \mathbf{u}_{k-1:k-N})$.

We augment the grey-box model in continuous time 
as follows -
\begin{align*}
\dot x_g &= f_g(x_g, u) + 
f_{NN}(\mathbf{y}_{k:k-N}, \mathbf{u}_{k:k-N})\\
y &= C_gx_g 
\end{align*}
in which $f_{NN}(\cdot)$,
is a standard feedforward neural network. 
We assume that the measurements
are directly some subset of the physical grey-box states.

It is expected that the function $f_{NN}(\cdot)$ approximates
the difference between the true grey-box state dynamics
with what is modeled. 
\begin{align*}
  f_{NN}(\mathbf{y}_{k:k-N}, \mathbf{u}_{k:k-N}) \approx 
      f_p(x, u) - f_g(x_g, u) = 
      Cf_p(\phi(\mathbf{y}_{k:k-N}, \mathbf{u}_{k-1:k-N}), u) - f_g(x_g, u)
\end{align*}

\subsection{Conversion to discrete time and training}

\subsection{Simple example}

%\textbf{Need to discuss the following issues:}
%\begin{itemize}
%  \item What are the example problems/applications 
%  where both kind of modeling 
%  strategies make sense, and where should we start.
%  Reproduce results from the Kevrekidis paper with measurement 
%  noise?
%  \item The model training step will be the optimization of 
%  the following multi-step-ahead prediction-error
%  \begin{equation*}
%    \underset{W_i, b_i}{\textnormal{min}} \sum_{k=0}^{t-1} (\hat{y}(k) - y(k))^2
%  \end{equation*}
%  in which ($W_i, b_i$) are the weights of the NN. This optimization 
%  is for one ID experiment, and requires an initial grey-box 
%  state $x_g(0)$ to make the forecast. If we have data 
%  for multiple ID experiments, we'll need initial states for 
%  all the ID experiments. In theory, we can add the initial 
%  states as decision variables in the optimization, but 
%  unfortunately tensorflow 
%  does not has a mechanism to do this.
%\end{itemize}

%\subsection{Adding feedforward disturbance forecasts}
%In this class of problems, we assume that we have a pretty 
%good u-y model from first-principles knowledge, and we'll use 
%neural networks to make feedforward disturbance forecasts
%for use in online MPC. HVAC is a pretty good application.

%\section{Optimization in MPC with neural networks as model}
%This is probably a late fall quarter or a winter break project. \\
%a) Try Casadi/MPC tools. 
%If tanh is the activation function (optimization is smooth),
%this should be fine. \\
%b) Try a global optimization solver. \\
%c) Train another NN with solutions from a global optimization solver.
